---
title: "Project Assignment 3"
author: "Elikem Asudo Tsatsu Gale-Zoyiku"
date: "2023-04-01"
output: word_document
---

```{r setup, message = FALSE}
```

## **Research Question**

**Is there an association between the sexual orientation of cis-males, depression levels, and their income? Can education explain any association if there is?**

Explanatory Variables: Gender (Categorical), Sexual Orientation (Categorical), and Educational attainment (categorical)

Response Variables: Depression levels (Categorical), Anxiety levels (Categorical), Ratio of income to poverty level (Categorical)

### 1. Load data set(s) and libraries

```{r 1}
load("C:/Users/egale/OneDrive - Ashesi University/Desktop/Statistics with Probability/IPUMS_NHIS.RData")
library(descr)
library(stats)
library(dplyr)
```

### 2. Create variable subset

```{r 2}
vars=c("AGE","SEX","SEXORIEN","EDUC","DEPFEELEVL","DEPFREQ","POVERTY","WORFEELEVL","WORFREQ")
myHealthData = IPUMS_NHIS[vars]
str(myHealthData)
myHealthData<-as.data.frame(myHealthData)
str(myHealthData)

```

### 3. Data management I: check for and recode errors and NAs

```{r 3}
myHealthData$Gender <- rep(NA,nrow(myHealthData))
myHealthData$Gender[myHealthData$SEX==1] <- 'Male'
myHealthData$Gender[myHealthData$SEX==2] <- 'Female'


myHealthData$WorkingAge <- rep(NA,nrow(myHealthData))
myHealthData$WorkingAge[myHealthData$AGE<=18] <- 'Not working age'
myHealthData$WorkingAge[myHealthData$AGE>18 & myHealthData$AGE<997] <- 'Working age'


myHealthData$Orientation <- rep(NA,nrow(myHealthData))
myHealthData$Orientation[myHealthData$SEXORIEN==1] <- 'Not Straight'
myHealthData$Orientation[myHealthData$SEXORIEN==3] <- 'Not Straight'
myHealthData$Orientation[myHealthData$SEXORIEN==4] <- 'Not Straight'
myHealthData$Orientation[myHealthData$SEXORIEN==2] <- 'Straight'

myHealthData$EDUC<-as.numeric(myHealthData$EDUC)
myHealthData$Education <- rep(NA, nrow(myHealthData))
myHealthData$Education[myHealthData$EDUC < 112] <- "No high school"
myHealthData$Education[myHealthData$EDUC >= 112 & myHealthData$EDUC<116] <- "Some high school"
myHealthData$Education[myHealthData$EDUC >= 200 & myHealthData$EDUC<=202] <- "High school diploma or equivalent"
myHealthData$Education[myHealthData$EDUC >= 300 & myHealthData$EDUC<=301] <- "Some college"
myHealthData$Education[myHealthData$EDUC >= 302 & myHealthData$EDUC<=303] <- "Associate's Degree"
myHealthData$Education[myHealthData$EDUC == 400] <- "Bachelor's Degree"
myHealthData$Education[myHealthData$EDUC == 504] <- "Other Degree"
myHealthData$Education[myHealthData$EDUC >= 500 & myHealthData$EDUC<=503] <- "Postgraduate"

myHealthData$DepressionFrequency <- rep(NA, nrow(myHealthData))
myHealthData$DepressionFrequency[myHealthData$DEPFREQ == 1] <- "Daily"
myHealthData$DepressionFrequency[myHealthData$DEPFREQ == 2] <- "Weekly"
myHealthData$DepressionFrequency[myHealthData$DEPFREQ == 3] <- "Monthly"
myHealthData$DepressionFrequency[myHealthData$DEPFREQ == 4] <- "Rarely"
myHealthData$DepressionFrequency[myHealthData$DEPFREQ == 5] <- "Never"


myHealthData$DepressionLevel <- rep(NA, nrow(myHealthData))
myHealthData$DepressionLevel[myHealthData$DEPFEELEVL == 1] <- "A lot"
myHealthData$DepressionLevel[myHealthData$DEPFEELEVL== 2] <- "A little"
myHealthData$DepressionLevel[myHealthData$DEPFEELEVL == 3] <- "Somewhere in between"
myHealthData$DepressionLevel[myHealthData$DEPFEELEVL == 0] <- NA


myHealthData$WorryFrequency <- rep(NA, nrow(myHealthData))
myHealthData$WorryFrequency[myHealthData$WORFREQ == 1] <- "Daily"
myHealthData$WorryFrequency[myHealthData$WORFREQ == 2] <- "Weekly"
myHealthData$WorryFrequency[myHealthData$WORFREQ == 3] <- "Monthly"
myHealthData$WorryFrequency[myHealthData$WORFREQ == 4] <- "Rarely"
myHealthData$WorryFrequency[myHealthData$WORFREQ == 5] <- "Never"


myHealthData$WorryLevel <- rep(NA, nrow(myHealthData))
myHealthData$WorryLevel[myHealthData$WORFEELEVL == 1] <- "A lot"
myHealthData$WorryLevel[myHealthData$WORFEELEVL== 2] <- "A little"
myHealthData$WorryLevel[myHealthData$WORFEELEVL == 3] <- "Somewhere in between"
myHealthData$WorryLevel[myHealthData$WORFEELEVL == 0] <- NA


myHealthData$PovertyRatio <- rep(NA, nrow(myHealthData))
myHealthData$PovertyRatio[myHealthData$POVERTY<=14] <- "Below the poverty line"
myHealthData$PovertyRatio[myHealthData$POVERTY<=25 & myHealthData$POVERTY>=20] <- "Between 1 to 1.99 units above the poverty line"
myHealthData$PovertyRatio[myHealthData$POVERTY<=32 & myHealthData$POVERTY>=31] <- "Between 2 to 2.99 units above the poverty line"
myHealthData$PovertyRatio[myHealthData$POVERTY<=34 & myHealthData$POVERTY>=33] <- "Between 3 to 3.99 units above the poverty line"
myHealthData$PovertyRatio[myHealthData$POVERTY<=36 & myHealthData$POVERTY>=35] <- "Between 4 to 4.99 units above the poverty line"
myHealthData$PovertyRatio[myHealthData$POVERTY==37] <- "5 units and above the poverty line"

```

### 4. Data management II: further subset and create secondary variable

```{r 4}
# Subset the data frame to take out observations below the working age
myHealthData <- myHealthData[myHealthData$WorkingAge == "Working age", ]

```

### 5. Descriptive statistics (sample means, standard deviations, proportions) and uni-variate displays

```{r 5}
freq(myHealthData$Gender, ylab="Frequency",xlab="Sex",ylim=c(0,70000),col=c("hotpink","blue"))
freq(myHealthData$WorkingAge, ylab="Frequency",xlab="Working Age",ylim=c(0,100000),col=c("pink","blue"))
freq(myHealthData$Orientation,ylab="Frequency",xlab="Sexual Orientation",ylim=c(0,100000), col = c("blue","yellow"))
freq(myHealthData$Education,ylab="Frequency",xlab="Highest Level of Education Obtained",ylim=c(0,50000),col=c("blue","yellow","green","violet","cyan","magenta"))
freq(myHealthData$DepressionFrequency,ylab="Frequency",xlab="How often people feel depressed",ylim=c(0,70000),col=c("blue","yellow","green","violet","cyan","magenta"))
freq(myHealthData$DepressionLevel,ylab="Frequency",xlab="How Severe is the depression, if depression is experienced",ylim=c(0,70000),col=c("blue","yellow","green","violet","cyan","magenta"))
freq(myHealthData$WorryFrequency,ylab="Frequency",xlab="How often people feel anxious",ylim=c(0,45000),col=c("blue","yellow","green","violet","cyan","magenta"))
freq(myHealthData$WorryLevel,ylab="Frequency",xlab="How severe is anxiety, if anxiety is experienced",ylim=c(0,60000),col=c("blue","yellow","green","violet","cyan","magenta"))
freq(myHealthData$PovertyRatio,ylab="Frequency",xlab="Ratio of income to poverty line",ylim=c(0,35000),col=c("blue","yellow","green","violet","cyan","magenta"))

t<-table(myHealthData$Gender)
p<-prop.table(t)*100;p

t<-table(myHealthData$WorkingAge)
p<-prop.table(t)*100;p

t<-table(myHealthData$Orientation)
p<-prop.table(t)*100;p

t<-table(myHealthData$Education)
p<-prop.table(t)*100;p

t<-table(myHealthData$DepressionFrequency)
p<-prop.table(t)*100;p

t<-table(myHealthData$DepressionLevel)
p<-prop.table(t)*100;p

t<-table(myHealthData$WorryFrequency)
p<-prop.table(t)*100;p

t<-table(myHealthData$WorryLevel)
p<-prop.table(t)*100;p

t<-table(myHealthData$PovertyRatio)
p<-prop.table(t)*100;p



```

### 6. Bivariate tables and graphs

```{r 6, fig.height=7.5, fig.width=10}
tab1<-table(myHealthData$PovertyRatio,myHealthData$Gender);tab1
prob_tab1<-round(prop.table(tab1,2),2)*100; prob_tab1
barplot(prob_tab1,main="Proportion of people in various levels above the poverty line by gender",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","black","gold"), beside = TRUE,legend=row.names(prob_tab1),ylim=c(0,70))

tab2<-table(myHealthData$PovertyRatio,myHealthData$Orientation);tab2
prob_tab2<-round(prop.table(tab2,2),2)*100;prob_tab2
barplot(prob_tab2,main="Proportion of people in various poverty levels by sexual orientation",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","black","gold"), beside = TRUE,legend=row.names(prob_tab2),ylim=c(0,70))

tab2<-table(myHealthData$Orientation,myHealthData$PovertyRatio);tab2
prob_tab2<-round(prop.table(tab2,2),2)*100
barplot(prob_tab2,main="Proportion of different sexual orientations in the various poverty levels",ylab="Proportion of Respondents", las=1, col=c(123,"pink"), beside = TRUE,legend=row.names(prob_tab2),ylim=c(0,100))

tab2<-table(myHealthData$PovertyRatio,myHealthData$Orientation);tab2
prob_tab2<-round(prop.table(tab2,2),2)*100
barplot(prob_tab2,main="Proportion of different sexual orientations in the various poverty levels",ylab="Proportion of Respondents", las=1, col=c(123,"pink"), beside = TRUE,legend=row.names(prob_tab2),ylim=c(0,100))

tab3<-table(myHealthData$Orientation,myHealthData$Gender);tab3
prob_tab3<-round(prop.table(tab3,2),2)*100
barplot(prob_tab3,main="Proportion of people's sexual orientation by gender",ylab="Proportion of Respondents", las=1, col=c(123,"pink"), beside = TRUE,legend=row.names(prob_tab3),ylim=c(0,100))


tab4 <- table(myHealthData$DepressionFrequency,myHealthData$Gender); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Proportion of depression frequency by gender",ylab="Proportion of Respondents", las=1, col=c("blue","red","green","black","gold"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))

tab4 <- table(myHealthData$DepressionLevel,myHealthData$Gender); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Proportion of depression level by gender",ylab="Proportion of Respondents", las=1, col=c("green","black","gold"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))

tab4 <- table(myHealthData$DepressionFrequency,myHealthData$Orientation); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Proportion of depression frequency by gender",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","beige"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))

tab4 <- table(myHealthData$DepressionLevel,myHealthData$Orientation); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Proportion of depression level by sexual orientation",ylab="Proportion of Respondents", las=1, col=c("pink",'red',"gold"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))

tab4 <- table(myHealthData$Education ,myHealthData$Gender); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Level of education by gender",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","beige","black"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))

tab4 <- table(myHealthData$Education ,myHealthData$Orientation); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Level of education by sexual orientation",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","beige","black"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))

tab4 <- table(myHealthData$PovertyRatio ,myHealthData$Education); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Poverty level by education level",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","black","gold"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))
```

I have learnt that the percentage of the population above working age that identifies as gay, bisexual. or lesbian is under 4%. This figure was unexpected. I also observed that more women than men identify as not straight. I also found that the proportion of people who identify as gay, lesbian or bisexual is fairly constant across all poverty levels. I observed that non-straight people seem to be have proportional or greater representation than straight people in the higher levels of education, such as postgraduate, bachelor's degree and some college education. Also, it is very evident that the higher the level of education, the higher the person's poverty level. These observations have interesting repercussions on my research questions.

### 7. Bivariate analysis (hypothesis tests and post-hoc tests)

```{r 7}
chisq_results <- chisq.test(myHealthData$DepressionFrequency, myHealthData$Gender)
chisq_results
chisq_results$observed
chisq_results$expected
chisq_results$residuals


chisq_results <- chisq.test(myHealthData$DepressionLevel, myHealthData$Gender)
chisq_results
chisq_results$observed
chisq_results$expected
chisq_results$residuals

chisq_results <- chisq.test(myHealthData$DepressionFrequency, myHealthData$Orientation)
chisq_results
chisq_results$observed
chisq_results$expected
chisq_results$residuals


chisq_results <- chisq.test(myHealthData$DepressionLevel, myHealthData$Orientation)
chisq_results
chisq_results$observed
chisq_results$expected
chisq_results$residuals

chisq_results <- chisq.test(myHealthData$PovertyRatio, myHealthData$Gender)
chisq_results
chisq_results$observed
chisq_results$expected
chisq_results$residuals

chisq_results <- chisq.test(myHealthData$PovertyRatio, myHealthData$Orientation)
chisq_results
chisq_results$observed
chisq_results$expected
chisq_results$residuals


chisq_results <- chisq.test(myHealthData$PovertyRatio ,myHealthData$Education)
chisq_results
chisq_results$observed
chisq_results$expected
chisq_results$residuals
stres <- chisq_results$stdres
```

To test the relationship between gender and depression levels which are both categorical variables, I conducted a Chi square test of independence. To conduct a Chi square test of independence, I followed these steps:

1.  I stated the null and alternative hypotheses:

    \
    Ho: There is no association between gender and depression levels.\
    H1: There is an association between gender and depression levels.

Summary of my samples:

|        |           |             |           |            |            |
|--------|-----------|-------------|-----------|------------|------------|
|        | Daily (%) | Monthly (%) | Never (%) | Rarely (%) | Weekly (%) |
| Female | 61        | 64          | 49        | 59         | 62         |
| Male   | 39        | 37          | 51        | 41         | 38         |

: Depression Frequency by Gender

|            |           |             |           |            |            |           |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
|            | **Daily** | **Monthly** | **Never** | **Rarely** | **Weekly** | **Total** |
| **Female** | 2287      | 4179        | 24215     | 14967      | 3274       | 48922     |
| **Male**   | 1462      | 2313        | 25150     | 10456      | 2037       | 41418     |
| **Total**  | 3749      | 6492        | 49365     | 25423      | 5311       | **90340** |

: Depression Frequency by Gender (Actual Counts)

|        |              |           |                          |
|--------|--------------|-----------|--------------------------|
|        | A little (%) | A lot (%) | Somewhere in between (%) |
| Female | 66           | 58        | 62                       |
| Male   | 34           | 42        | 38                       |

: Depression Level of respondents by Gender

|            |              |           |                          |           |
|------------|--------------|-----------|--------------------------|-----------|
|            | **A little** | **A lot** | **Somewhere in between** | **Total** |
| **Female** | 3498         | 12381     | 9359                     | 25238     |
| **Male**   | 1839         | 8865      | 5765                     | 16469     |
| **Total**  | 5337         | 21246     | 15124                    | **41707** |

: Depression Level of respondents by Gender (Actual Counts)

//the counts in this table are less because I took out the respondents who said they had never felt depressed

The conditions for a Chi-Squared test are met because:

-   Both gender and depression frequency / depression level are categorical variables.

-   The cells in the contingency table are mutually exclusive, i.e. nobody reports that they experience two or more levels of depression, and nobody reports that they are both male and female.

-   The expected value of all the cells are greater than 5, and no cell has an expected value less than 1.

Test Statistic (Chi-Square):

-   

    ```         
    Pearson's Chi-squared test

    data:  myHealthData$DepressionFrequency and myHealthData$Gender
    X-squared = 1209.2, df = 4, p-value < 2.2e-16
    ```

-   

    ```         
    Pearson's Chi-squared test

    data:  myHealthData$DepressionLevel and myHealthData$Gender
    X-squared = 112.91, df = 2, p-value < 2.2e-16
    ```

-   The p-value of the test for the association between gender and depression frequency is significantly less than 0.05, so I have sufficient evidence to reject my null hypothesis. This implies that there is an association between a person's gender and how frequently they experience depression.

-   The p-value of the test for the association between gender and depression level is significantly less than 0.05, so I have sufficient evidence to reject my null hypothesis. This implies that there is an association between a person's gender and how severe the person's depression is.

-   The type of error that could have been made in this test is a type I error, which occurs when the null hypothesis is rejected when it is actually true. This means that I could have falsely concluded that there is an association between gender and depression when there is none. The probability of making a type I error is equal to the significance level (alpha) for this test (0.05).

I repeated this test four (4) more times for different sets of hypothesis. These are:

**Set 2:**\

To test the relationship between sexual orientation and depression levels which are both categorical variables, I conducted a Chi square test of independence. To conduct a Chi square test of independence, I followed these steps:

I stated the null and alternative hypotheses:

Ho: There is no association between sexual orientation and depression levels.

H1: There is an association between sexual orientation and depression levels.

Summary of my samples:

```         
Frequency of depression by sexual orientation (actual count)

              Not Straight Straight
  Daily            313     3281
  Monthly          493     5831
  Never           1006    47099
  Rarely          1079    23772
  Weekly           513     4634
```

```         
Level of depression by sexual orientation (actual count)

                        Not Straight Straight
  A little                      530     4632
  A lot                         845    19929
  Somewhere in between         1056    13653
```

```         
Frequency of depression by sexual orientation (proportions)           
           
            Not Straight Straight
  Daily              9        4
  Monthly           14        7
  Never             30       56
  Rarely            32       28
  Weekly            15        5
```

```         
Level of depression by sexual orientation (proportions)      
      
                    
                       Not Straight Straight
  A little                       22       12
  A lot                          35       52
  Somewhere in between           43       36
```

The conditions for a Chi-Squared test are met because:

-   Both sexual orientation and depression frequency / depression level are categorical variables.

-   The cells in the contingency table are mutually exclusive, i.e. nobody reports that they experience two or more levels of depression, and nobody reports that they are both straight and not straight.

-   The expected value of all the cells are greater than 5, and no cell has an expected value less than 1.

Test Statistic (Chi-Square):

-   

```         
Pearson's Chi-squared test

data:  myHealthData$DepressionFrequency and myHealthData$Orientation
X-squared = 1427.4, df = 4, p-value < 2.2e-16
```

-   

    ```         
    Pearson's Chi-squared test

    data:  myHealthData$DepressionLevel and myHealthData$Orientation
    X-squared = 341.47, df = 2, p-value < 2.2e-16
    ```

-   The p-value of the test of the association between sexual orientation and depression frequency is significantly less than 0.05, so I have sufficient evidence to reject my null hypothesis. This implies that there is an association between a person's sexual orientation and how frequently they experience depression.

-   The p-value of the test for the association between sexual orientation and depression level is significantly less than 0.05, so I have sufficient evidence to reject my null hypothesis. This implies that there is an association between a person's sexual orientation and how severe the person's depression is.

-   The type of error that could have been made in this test is a type I error, which occurs when the null hypothesis is rejected when it is actually true. This means that I could have falsely concluded that there is an association between sexual orientation and depression when there is none. The probability of making a type I error is equal to the significance level (alpha) for this test (0.05).

**Set 3:**

\
To test the relationship between gender and poverty which are both categorical variables, I conducted a Chi square test of independence. To conduct a Chi square test of independence, I followed these steps:

I stated the null and alternative hypotheses:

Ho: There is no association between gender and poverty level.\
H1: There is an association between gender and poverty level.

Summary of my samples:

```         
  Distribution of different genders across different poverty levels (Actual counts)
  
                                                  Female  Male
  5 units and above the poverty line              14400 15097
  Below the poverty line                           5768  3369
  Between 1 to 1.99 units above the poverty line   9500  6450
  Between 2 to 2.99 units above the poverty line   8286  6545
  Between 3 to 3.99 units above the poverty line   6492  5820
  Between 4 to 4.99 units above the poverty line   5486  4916
  
  
  Distribution of different genders across different poverty levels (proportions)
  
                                                  Female Male
  5 units and above the poverty line                 29   36
  Below the poverty line                             12    8
  Between 1 to 1.99 units above the poverty line     19   15
  Between 2 to 2.99 units above the poverty line     17   16
  Between 3 to 3.99 units above the poverty line     13   14
  Between 4 to 4.99 units above the poverty line     11   12
```

The conditions for a Chi-Squared test are met because:

-   Both poverty level and gender are categorical variables.

-   The cells in the contingency table are mutually exclusive, i.e. nobody is in more than one poverty level, and nobody reports that they are both male and female.

-   The expected value of all the cells are greater than 5, and no cell has an expected value less than 1.

Test Statistic (Chi-Square):

```         
    Pearson's Chi-squared test

data:  myHealthData$PovertyRatio and myHealthData$Gender
X-squared = 858.5, df = 5, p-value < 2.2e-16
```

The p-value of the test of the association between poverty level and gender is significantly less than 0.05, so I have sufficient evidence to reject my null hypothesis. This implies that there is an association between a person's poverty level and their gender.

The type of error that could have been made in this test is a type I error, which occurs when the null hypothesis is rejected when it is actually true. This means that I could have falsely concluded that there is an association between gender and poverty ratio when there is none. The probability of making a type I error is equal to the significance level (alpha) for this test (0.05).

**Set 4:**\

To test the relationship between sexual orientation and poverty which are both categorical variables, I conducted a Chi square test of independence. To conduct a Chi square test of independence, I followed these steps:

I stated the null and alternative hypotheses:

Ho: There is no association between sexual orientation and poverty level.\
H1: There is no association between sexual orientation and poverty level.

Summary of my samples:

```         
:Distribution of different sexual orientations across different poverty levels (Actual counts)

                                                  Not Straight Straight
  5 units and above the poverty line                      1064    27477
  Below the poverty line                                  392     8214
  Between 1 to 1.99 units above the poverty line          590    14637
  Between 2 to 2.99 units above the poverty line          571    13630
  Between 3 to 3.99 units above the poverty line          440    11434
  Between 4 to 4.99 units above the poverty line          361     9740
```

```         
:Distribution of different sexual orientations across different poverty levels (proportions)                                              
                                                 Not Straight Straight
  5 units and above the poverty line                       31       32
  Below the poverty line                                   11       10
  Between 1 to 1.99 units above the poverty line           17       17
  Between 2 to 2.99 units above the poverty line           17       16
  Between 3 to 3.99 units above the poverty line           13       13
  Between 4 to 4.99 units above the poverty line           11       11
```

The conditions for a Chi-Squared test are met because:

-   Both poverty level and sexual orientation are categorical variables.

-   The cells in the contingency table are mutually exclusive, i.e. nobody is in more than one poverty level, and nobody reports that they are both male and female.

-   The expected value of all the cells are greater than 5, and no cell has an expected value less than 1.

Test Statistic (Chi-Square):

```         
Pearson's Chi-squared test

data:  myHealthData$PovertyRatio and myHealthData$Orientation
X-squared = 16.531, df = 5, p-value = 0.005481
```

The p-value of the test of the association between poverty level and sexual orientation is significantly less than 0.05, so I have sufficient evidence to reject my null hypothesis. This implies that there is an association between a person's poverty level and their sexual orientation.

The type of error that could have been made in this test is a type I error, which occurs when the null hypothesis is rejected when it is actually true. This means that I could have falsely concluded that there is an association between sexual orientation and poverty ratio when there is none. The probability of making a type I error is equal to the significance level (alpha) for this test (0.05).

### 8. Moderation

```{r 8, fig.height=10, fig.width=20}

tab4 <- table(myHealthData$Education ,myHealthData$Gender); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Level of education by gender",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","beige","black"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))

tab4 <- table(myHealthData$Education ,myHealthData$Orientation); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Level of education by sexual orientation",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","beige","black"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))

tab4 <- table(myHealthData$PovertyRatio ,myHealthData$Education); tab4
prop_tab4 <- round(prop.table(tab4,2),2)*100; prop_tab4
barplot(prop_tab4,main="Poverty level by education level",ylab="Proportion of Respondents", las=1, col=c("pink","blue","red","green","black","gold"), beside = TRUE,legend=row.names(prop_tab4),ylim=c(0,70))
```

### 9. Save

```{r 9}
save(myHealthData, file = "myHealthData.RDATA")
```
