---
title: "Project Assignment 3"
author: "Elikem Asudo Tsatsu Gale-Zoyiku"
date: "2023-04-01"
output: word_document
---
```{r setup, message = FALSE}
```

## **Research Questions**

**What is the relationship between exercise frequency and mental health status?**

Explanatory Variable: Exercise Frequency (Quantitative)
Response Variable: Mental Health Status (Categorical)

### 1. Load data set(s) and libraries
```{r 1}
load("C:/Users/egale/OneDrive - Ashesi University/Desktop/Statistics with Probability/IPUMS_NHIS.RData")
library(descr)
library(stats)
```


### 2. Create variable subset
```{r 2}
vars=c("SEX","SEXORIEN","EDUC","MAXEDUC","WORFREQ","DEPFEELEVL","WORRX","DEPRX","DEPFREQ","POVLEV","POVERTY")
myData = IPUMS_NHIS[vars]

str(myData)
myData<-as.data.frame(myData)
str(myData)

```

### 3. Data management I: check for and recode errors and NAs
```{r 3}
```

### 4. Data management II: further subset and create secondary variable
```{r 4}
```

### 5. Descriptive statistics (sample means, standard deviations, proportions) and univariate displays
```{r 5}
```

### 6. Bivariate tables and graphs
```{r 6}
```

### 7. Bivariate analysis (hypothesis tests and post-hoc tests)
```{r 7}
```

### 8. Moderation
```{r 8}
```

### 9. Save
```{r 9}
```